# Reproducible Research: Peer Assessment 1
================================================================
Submitted by Lorena Zuniga

In this asignment 


## Loading and preprocessing the data
Data from activity monitoring device, ignoring NA values.
```{r, echo=TRUE}
  activityData_all <- read.csv('activity.csv')
  
  activityData <- na.omit(activityData_all)

  
```


## What is mean total number of steps taken per day?

Histogram of the total number of steps taken each day

```{r histogram, echo=TRUE}
    
    activityData$date<-as.Date(activityData$date)
   
    totalByDay <- aggregate(activityData$steps,list(activityData$date),FUN=sum)
    hist(totalByDay$x,main='Total steps taken each day',xlab='Total steps')

```

### The mean and median total number of steps taken per day:
```{r, echo=TRUE}
 
   theMean <-mean(totalByDay$x)
   theMedian <- median(totalByDay$x)
  
 
```
The mean is `r theMean`  and the median is `r theMedian`


## What is the average daily activity pattern?
```{r, echo=TRUE}
  activityPattern <-aggregate(activityData$steps,list(activityData$interval),FUN=mean)
  
  colnames(activityPattern) <- c('Interval','Average')
  
  plot(activityPattern$Interval,activityPattern$Average,type='l',xlab='Interval',ylab='Average Steps', main='Average steps taken by 5-minute interval')
```

## Imputing missing values
### Calculating the total number of missing values in the dataset
```{r, echo=TRUE}
   totalMissing= sum(is.na(activityData_all))
```
Number of missing values is `r totalMissing`

### Substitution of NA values
NA values are substituted by the mean of the correspondent 5minute interval

```{r,echo=TRUE}
  
  for(i in 1:nrow(activityData_all))
    {
       if (is.na(activityData_all[i,]$steps))
       {
         curInterval <- activityData_all[i,]$interval
         meanSteps <- activityPattern[activityPattern$Interval==curInterval,]$Average
         activityData_all[i,]$steps<-meanSteps
       }
    }
 
```

### New Total number of steps taken each day (with replaced NA values)
Histogram of the total number of steps taken each day
```{r histogram2, echo=TRUE}
    activityData_all$date <- as.Date(activityData_all$date)

    newTotalByDay <- aggregate(activityData_all$steps,list(activityData_all$date),FUN=sum)
    hist(newTotalByDay$x, main='Total steps taken each day', xlab='Total steps')
```

### The mean and median total number of steps taken per day:
```{r, echo=TRUE}
 
   theNewMean <-mean(newTotalByDay$x)
   theNewMedian <- median(newTotalByDay$x)
  
```
The mean is `r theNewMean`  and the median is `r theNewMedian`

In this case the mean and median obtained are practically the same as before the NA values substitution.

## Are there differences in activity patterns between weekdays and weekends?

Adding a new column (typeOfDay) to the dataframe in order to indicate if the day is a weekday or a weekend day
```{r, echo=TRUE}
    activityData_all$typeOfDay <- ifelse(weekdays(activityData_all$date)== 'Saturday' | weekdays (activityData_all$date)== 'Sunday','weekend','weekday')
```

### Plotting the average number of steps taken (y axis) on weekdays or weekend days (x-axis)

